---
title: "Compare worm and mouse DTU"
author: "Emma Jones"
date: "2023-07-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

For code review, I am tracking the amount of time that this analysis will take.

```{r proc time}
ptm <- proc.time()
```

The purpose of this script is to compare DTU between mouse and worm. It is dependent on script 17 and scripts 01-07.

```{r load packages and set seed}
suppressPackageStartupMessages({
  library(tidyverse)
  library(IsoformSwitchAnalyzeR)
  library(ComplexHeatmap)
  library(viridis)
  library(org.Ce.eg.db)
  library(gprofiler2)
  library(styler)
  library(lintr)
  library(here)
})

set.seed(123)
```



```{r read in data}
worm_switchlist_analyzed <- readRDS("/data/user/efjones/230227_EJ_MouseBrainIsoDiv/data/switchlist_objects/raw/worm_switchlist_analyzed.Rds")

region_sex_switchlist_list <- readRDS("/data/user/efjones/230227_EJ_MouseBrainIsoDiv/data/switchlist_objects/raw/region_sex_switchlist_list.Rds")

sex_switchlist <- readRDS("/data/user/efjones/230227_EJ_MouseBrainIsoDiv/data/switchlist_objects/raw/sex_switchlist.Rds")

```

Need to filter for significant genes

```{r }
cerebellum_cortex_list <- list(
  `worm sex DTU` = cerebellum_cortex_dge_genes,
  `mouse sex DTU` = cerebellum_cortex_dte_genes,
  `mouse cerebellum sex DTU` = cerebellum_cortex_dtu_genes,
  `mouse cortex sex DTU` = cerebellum_cortex_dtu_genes,
  `mouse striatum sex DTU` = cerebellum_cortex_dtu_genes
)
```

```{r proc time finish}
fptm <- proc.time() - ptm
fptm[3] / 60
```

#### Clean up script

```{r tidy script}
style_file("17_dtu_whole_worm.Rmd")

lint("17_dtu_whole_worm.Rmd",
  linters = linters_with_defaults(
    object_length_linter = NULL,
    object_name_linter = NULL,
    object_usage_linter = NULL
  )
)
```


```{r versions}
sessionInfo()
```
R version 4.3.0 beta (2023-04-12 r84240)
Platform: x86_64-pc-linux-gnu (64-bit)
Running under: Ubuntu 22.04.2 LTS

Matrix products: default
BLAS:   /usr/lib/x86_64-linux-gnu/openblas-pthread/libblas.so.3 
LAPACK: /usr/lib/x86_64-linux-gnu/openblas-pthread/libopenblasp-r0.3.20.so;  LAPACK version 3.10.0

locale:
 [1] LC_CTYPE=en_US.UTF-8       LC_NUMERIC=C               LC_TIME=en_US.UTF-8       
 [4] LC_COLLATE=en_US.UTF-8     LC_MONETARY=en_US.UTF-8    LC_MESSAGES=en_US.UTF-8   
 [7] LC_PAPER=en_US.UTF-8       LC_NAME=C                  LC_ADDRESS=C              
[10] LC_TELEPHONE=C             LC_MEASUREMENT=en_US.UTF-8 LC_IDENTIFICATION=C       

time zone: Etc/UTC
tzcode source: system (glibc)

attached base packages:
[1] grid      stats4    stats     graphics  grDevices utils     datasets  methods  
[9] base     

other attached packages:
 [1] here_1.0.1                    lintr_3.0.2                  
 [3] styler_1.9.1                  gprofiler2_0.2.1             
 [5] org.Ce.eg.db_3.17.0           viridis_0.6.2                
 [7] viridisLite_0.4.2             ComplexHeatmap_2.16.0        
 [9] IsoformSwitchAnalyzeR_1.99.17 pfamAnalyzeR_1.0.0           
[11] sva_3.48.0                    genefilter_1.82.0            
[13] mgcv_1.8-42                   nlme_3.1-162                 
[15] satuRn_1.8.0                  DEXSeq_1.46.0                
[17] RColorBrewer_1.1-3            AnnotationDbi_1.62.0         
[19] DESeq2_1.40.0                 SummarizedExperiment_1.30.0  
[21] GenomicRanges_1.52.0          GenomeInfoDb_1.36.0          
[23] IRanges_2.34.0                S4Vectors_0.38.0             
[25] MatrixGenerics_1.12.0         matrixStats_0.63.0           
[27] Biobase_2.60.0                BiocGenerics_0.46.0          
[29] BiocParallel_1.34.0           limma_3.56.0                 
[31] lubridate_1.9.2               forcats_1.0.0                
[33] stringr_1.5.0                 dplyr_1.1.2                  
[35] purrr_1.0.1                   readr_2.1.4                  
[37] tidyr_1.3.0                   tibble_3.2.1                 
[39] ggplot2_3.4.2                 tidyverse_2.0.0              

loaded via a namespace (and not attached):
  [1] splines_4.3.0                 later_1.3.0                  
  [3] BiocIO_1.10.0                 bitops_1.0-7                 
  [5] filelock_1.0.2                R.oo_1.25.0                  
  [7] rex_1.2.1                     XML_3.99-0.14                
  [9] lifecycle_1.0.3               rprojroot_2.0.3              
 [11] edgeR_3.42.0                  doParallel_1.0.17            
 [13] processx_3.8.0                lattice_0.21-8               
 [15] ensembldb_2.24.0              magrittr_2.0.3               
 [17] plotly_4.10.1                 rmarkdown_2.21               
 [19] remotes_2.4.2                 yaml_2.3.7                   
 [21] httpuv_1.6.9                  pbapply_1.7-0                
 [23] DBI_1.1.3                     zlibbioc_1.46.0              
 [25] R.cache_0.16.0                R.utils_2.12.2               
 [27] AnnotationFilter_1.24.0       RCurl_1.98-1.12              
 [29] rappdirs_0.3.3                circlize_0.4.15              
 [31] GenomeInfoDbData_1.2.10       annotate_1.78.0              
 [33] codetools_0.2-19              DelayedArray_0.25.0          
 [35] xml2_1.3.4                    tidyselect_1.2.0             
 [37] shape_1.4.6                   futile.logger_1.4.3          
 [39] locfdr_1.1-8                  BiocFileCache_2.8.0          
 [41] GenomicAlignments_1.36.0      jsonlite_1.8.4               
 [43] GetoptLong_1.0.5              ellipsis_0.3.2               
 [45] survival_3.5-5                iterators_1.0.14             
 [47] foreach_1.5.2                 tools_4.3.0                  
 [49] progress_1.2.2                Rcpp_1.0.10                  
 [51] glue_1.6.2                    gridExtra_2.3                
 [53] xfun_0.39                     withr_2.5.0                  
 [55] formatR_1.14                  BiocManager_1.30.20          
 [57] fastmap_1.1.1                 boot_1.3-28.1                
 [59] fansi_1.0.4                   callr_3.7.3                  
 [61] digest_0.6.31                 timechange_0.2.0             
 [63] R6_2.5.1                      mime_0.12                    
 [65] colorspace_2.1-0              biomaRt_2.56.0               
 [67] RSQLite_2.3.1                 R.methodsS3_1.8.2            
 [69] utf8_1.2.3                    generics_0.1.3               
 [71] data.table_1.14.8             tximeta_1.18.0               
 [73] rtracklayer_1.60.0            htmlwidgets_1.6.2            
 [75] prettyunits_1.1.1             httr_1.4.5                   
 [77] pkgconfig_2.0.3               gtable_0.3.3                 
 [79] blob_1.2.4                    hwriter_1.3.2.1              
 [81] XVector_0.40.0                htmltools_0.5.5              
 [83] geneplotter_1.78.0            ProtGenerics_1.32.0          
 [85] clue_0.3-64                   scales_1.2.1                 
 [87] cyclocomp_1.1.0               png_0.1-8                    
 [89] knitr_1.42                    lambda.r_1.2.4               
 [91] rstudioapi_0.14               tzdb_0.3.0                   
 [93] reshape2_1.4.4                rjson_0.2.21                 
 [95] curl_5.0.0                    cachem_1.0.8                 
 [97] GlobalOptions_0.1.2           BiocVersion_3.17.1           
 [99] parallel_4.3.0                desc_1.4.2                   
[101] restfulr_0.0.15               pillar_1.9.0                 
[103] vctrs_0.6.3                   promises_1.2.0.1             
[105] dbplyr_2.3.2                  xtable_1.8-4                 
[107] cluster_2.1.4                 tximport_1.28.0              
[109] evaluate_0.20                 VennDiagram_1.7.3            
[111] GenomicFeatures_1.52.0        cli_3.6.1                    
[113] locfit_1.5-9.7                compiler_4.3.0               
[115] futile.options_1.0.1          Rsamtools_2.16.0             
[117] rlang_1.1.1                   crayon_1.5.2                 
[119] ps_1.7.4                      plyr_1.8.8                   
[121] stringi_1.7.12                munsell_0.5.0                
[123] Biostrings_2.68.0             lazyeval_0.2.2               
[125] Matrix_1.5-4                  BSgenome_1.68.0              
[127] hms_1.1.3                     bit64_4.0.5                  
[129] KEGGREST_1.40.0               statmod_1.5.0                
[131] shiny_1.7.4                   interactiveDisplayBase_1.38.0
[133] AnnotationHub_3.8.0           memoise_2.0.1                
[135] bit_4.0.5 